---
title: "P8105 Homework 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem 1

```{r creating dataframe}

set.seed(1)

df_p1 <- tibble(
  vec_numeric = runif(10, 0, 5), 
  vec_logical = vec_numeric > 2, 
  vec_char = c("a", "b", "c", "d", "e", "f", "g", "h", "j", "k"), 
  vec_factor = factor(c("male", "female", "male", "female", "female", "male", "male", "female", "female", "male"))
)

```

```{r means}

mean(df_p1$vec_numeric)
mean(df_p1$vec_logical)
mean(df_p1$vec_char)
mean(df_p1$vec_factor)

```

I am able to take the mean of the numeric and logical vector. However, I cannot take the mean of the character or factor vector; when I try an error is returned indicating the argument is not numeric or logical. I'm assuming I can take the mean of a logical vector because R internally represents false as 0 and true as 1. 

```{r converting practice 1, eval = FALSE}

as.numeric(df_p1$vec_logical)
as.numeric(df_p1$vec_char)
as.numeric(df_p1$vec_factor)

```

When I convert the logical vector to numeric a vector of 0 and 1s is returned, confirming my suspicion about R's internal representation of logical values. For the character vector, NAs are returned when I try to convert the vector to numeric. When I convert the factor vector to numeric a vector of 1s and 2s are returned. I believe these correspond to the 1st and the 2nd level of the factor respectively. 

```{r converting practice 2, eval = FALSE}

as.numeric(as.factor(df_p1$vec_char))
as.numeric(as.character(df_p1$vec_factor))

```

When I convert the character vector to a factor and then to a numeric vector a vector of values ranging from 1:10 are returned. This represent R first factoring the 10 character values and then returning the levels of the factor when coerced to numeric. NAs are returned when I convert the factor vector to character and then to numeric. This is becuase R is unable to convert character values to numeric. 






